visualFields 1.0.7: Tools for Analyzing the Field of Vision in \proglang{R}

### 2025-05-18 jss-article-rnw/article.Rnw
Incremental improvement of manuscript.

%% another package (only for this demo article)
\usepackage{framed}
\usepackage{listings}

@@

\begin{document}
%%\SweaveOpts{concordance=TRUE}
\lstset{breaklines=true, breakatwhitespace=true}

@@

\subsection{False negatives and false positives}

@@

@@

\subsection{Working with visual field data}

Each row represents a monocular test. The eye is labeled using Latin abbreviations:

@@

@@

\subsection{Metric for individual locations}

@@

<<>>=
td <- gettd(example2)
print(td[, getvfcols()[1:10]])
@

@@

<<>>=
pd <- getpd(td)
print(pd[, getvfcols()[1:10]])
@

@@

<<>>=
print(getnv()$lut$probs)
print(gettdp(td)[, getvfcols()[1:10]])
print(getpdp(pd)[, getvfcols()[1:10]])
@

@@

<<>>=
prog1 <- plr(example3)
str(prog1, max.level = 1, strict.width = "wrap")
print(sum(prog1$pval[-getlocmap()$bs] < 0.05))
@

@@

<<>>=
print(prog2$csr)
print(prog2$csrp)
print(prog2$csl)
print(prog2$cslp)
@

@@

<<>>=
hist(prog2$cstats$cslall, xlab = "S-values", xlim = c(0, 200))
abline(v = prog2$csl, col = "red", lwd = 2)
@

@@

\section{More technical details} \label{app:technical}

\begin{leftbar}
Appendices can be included after the bibliography (with a page break). Each
section within the appendix should have a proper section title (rather than
just \emph{Appendix}).

For more technical style details, please check out JSS's style FAQ at
\url{https://www.jstatsoft.org/pages/view/style#frequently-asked-questions}
which includes the following topics:
\begin{itemize}
  \item Title vs.\ sentence case.
  \item Graphics formatting.
  \item Naming conventions.
  \item Turning JSS manuscripts into \proglang{R} package vignettes.
  \item Trouble shooting.
  \item Many other potentially helpful details\dots
\end{itemize}
\end{leftbar}

>>>>

%% another package (only for this demo article)
\usepackage{framed}

@@

\begin{document}
%%\SweaveOpts{concordance=TRUE}
%\lstset{breaklines=true, breakatwhitespace=true}

@@

For analyzing visual field data, the package provides the following functionality:
\begin{itemize}
  \item A format for representing visual field data
  \item Functions for importing data from perimeters
  \item Functions for reading and writing data in CSV format, and for handlings data
  \item A format for representing the underlying spatial pattern ({data(locmaps)})
  \item A format for representing age-corrected normal values ({data(normvals)}), including probabilities
  \item Functions for calculating total deviation (TD), pattern deviation (PD), and global indices
  \item Functions for calculating probabilities of TD and PD
  \item A format for representing parameters for plotting the visual fields ({data(gpars)})
  \item Functions for plotting visual fields, including longitudinal analyses of visual field series
  \item Functions for calculating structural parameters pertinent to visual field testing (Jansonius maps, ganglion cell displacement)
\end{itemize}

By default, the package is set up to handle Humphrey Field Analyzer 24-2 visual fields. It is recommended to use such data for exploring the package and data is provided by the package. However, other patterns and devices can be used. The package contains all necessary data for the HFA 10-2 pattern. The package is designed to be extensible, allowing users to add new patterns, devices, and normative datasets.

@@

\subsection{Visual field data}

The first ten columns contain metadata for each test, including reliability indices (false negatives, false positives, and fixation losses). The remaining columns contain the actual test values (sensitivities) at the locations defined in the \code{locmap}.

The \code{date} variable is stored as a \code{POSIXct} object, which can be converted to a date format using the \code{as.Date()} function. Each row represents a monocular test.

The eye is labeled using Latin abbreviations:

The \code{type} variable allows assignment of tests to different groups.

\subsubsection{Reliability indices}

\subsubsection{Psychophysical data: sensitivities}

@@

@@

\subsection{Graphical parameters}

<<>>=
data(gpars)
names(gpars)
str(gpars$p24d2, 
    max.level = 1, strict.width = "wrap")
@

The \code{gpars} object contains graphical parameters for plotting visual fields. The default parameters are set for the 24-2 pattern, but can be customized for other patterns. It contains a list of the coords, which must be identical to the \code{locmap} used for the test. The \code{gpar} object also contains information on how to tesselate the borders in the plots, and color palettes for both the fields and their borders in the visual field plots.

@@

\subsection{Metrics for individual locations}

@@

<<>>=
td <- gettd(example2)
td[, getvfcols()[1:10]]
@

@@

<<>>=
pd <- getpd(td)
pd[, getvfcols()[1:10]]
@

@@

<<>>=
getnv()$lut$probs
gettdp(td)[, getvfcols()[1:10]]
getpdp(pd)[, getvfcols()[1:10]]
@

@@

<<>>=
prog1 <- plr(example3)
str(prog1, max.level = 1, strict.width = "wrap")
sum(prog1$pval[-getlocmap()$bs] < 0.05)
@

@@

<<>>=
summary(prog2)
prog2[c("csr", "csrp", "csl", "cslp")]
@

@@

<<fig=TRUE>>=
ggplot(data.frame(x = prog2$cstats$cslall), aes(x = x)) + 
         geom_histogram() +
         geom_vline(xintercept = prog2$csl, color = "red", size = 1) +
         scale_x_continuous("S-values", breaks = seq(0, 200, 20))
@

@@

\section{Clinical aspects} \label{app:clinical}
